News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-09-04 rupnik2015.tex
updates

The paper is organized as follows: we first provide an overview of the system as a whole (Section ~\ref{sec:pipeline}), followed by details on each of the two steps described above. We first introduce the problem of cross-lingual similarity computation (Section~\ref{sec:crosslingual}), which is followed by several approaches for measuring cross-lingual similarity (Section~\ref{sec:models}) and then the algorithms for matching clusters in different languages (Section~\ref{sec:linking}).  To improve readability, we describe relevant related work for each module in its corresponding section. Finally, we present and interpret the experimental results, give conclusions and describe promising future directions.

@@

We base our techniques of cross-lingual event linking on an online system for detection of world events, called Event Registry \shortcite{Leban2014W,Leban2014I}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles that are collected from numerous news outlets all over the world. The important components in the pipeline of the Event Registry are shown in Figure~\ref{fig:erpipeline} and are briefly described below.

@@

\subsection{Cross-Lingual Models}\label{sec:models}
In this section, we describe several approaches to the problem of computing multilingual simliarities introduced in \ref{sec:tfidf}. We present four approaches:
a simple approach based on $k$-means clustering in \ref{sec:kmeans}, a standard approach based on singular value decomposition in \ref{sec:LSI}, a related
approach called Canonical Correlation Analysis (CCA) in \ref{sec:CCA} and finally a new method, which is an extension of CCA to more than two languages in \ref{sec:hublang}.
%
CCA can be used to find correlated patterns for a pair of languages, whereas the extended method optimizes a
Sum of Squared Correlations (SSCOR) between several language pairs, which was introduced in \cite{Kettenring}. The SSCOR problem is difficult to solve in our setting (hundreds of thousands of features, hundreds of thousands of examples). To tackle this, we propose a method which consists of two ingredients.
 The first one is based on an observation that certain datasets (such as Wikipedia) are biased towards one language (English), which can be exploited
 to reformulate a difficult optimization problem as an eigenvector problem. The second ingredient is dimensionality reduction using LSI, which
 makes the eigenvector problem computationally and numerically tractable.

@@

The resulting matrix (when appropriately scaled to have unit norm) for similarity computation between language $i$ and language $j$ is defined as (see the reasoning below):
$$C_i(C_i^T C_i)^{-1} (C_j^T C_j)^{-1} C_j.$$

@@

The first step in our method is to project $X_1, \ldots, X_m$ to lower-dimensional spaces without destroying the cross-lingual structure. Treating the nonzero columns of $X_i$ as observation vectors sampled from an underlying distribution $\mathcal{X}_i \in V_i = \RR^{n_i}$, we can analyze the empirical cross-covariance matrices:
$$C_{i,j} = \frac{1}{|a(i,j)|-1 }\sum_{\ell \in a(i,j)} (X_i^{\ell} - c_i)\cdot (X_j^{\ell} - c_j)^T,$$
 where $c_i = \frac{1}{a_i} \sum_{\ell \in a(i)}X_i^{\ell}$. By finding low-rank approximations of $C_{i,j}$ we can identify the subspaces of $V_i$ and $V_j$ that are relevant for extracting linear patterns between $\mathcal{X}_i$ and $\mathcal{X}_j$. Let $X_1$ represent the hub language corpus matrix. The LSI approach to finding the subspaces is to perform the singular value decomposition on the full $N \times N$ covariance matrix composed of blocks $C_{i,j}$. If $|a(i,j)|$ is small for many language pairs (as it is in the case of Wikipedia), then many empirical estimates $C_{i,j}$ are unreliable, which can result in overfitting. For this reason, we perform the truncated singular value decomposition on the matrix $C = [C_{1,2}  \cdots  C_{1,m}] \approx U S V^T$, where $U \in \RR^{n_1 \times k}, S \in \RR^{k \times k}, V \in \RR^{(\sum_{i=2}^m n_i) \times k}$. We split the matrix $V$ vertically in blocks with $n_2, \ldots, n_m$ rows: $V = [V_2^T  \cdots  V_m^T]^T$. Note that columns of $U$ are orthogonal but columns in each $V_i$ are not (columns of V are orthogonal). Let $V_1 := U$. We proceed by reducing the dimensionality of each $X_i$ by setting: $Y_i = V_i^T \cdot X_i$, where $Y_i \in \RR^{k\times N}$. To summarize, the first step reduces the dimensionality of the data and is based on CL-LSI, but optimizes only the hub-language related cross-covariance blocks.

@@

We describe the main dataset for building cross-lingual models which is based on Wikipedia and then present two sets of experiments. The first set of experiments
establishes that the hub based approach can deal with language pairs where little or no training data is available. The second set of experiments compares the main approaches
that we presented on the task of mate retrieval and the task of event linking. Finally, we examine how different choices of features impact the event linking performance.

@@

To investigate the empirical performance of the low-rank approximations we will test the algorithms on a large-scale, real-world multilingual dataset that we extracted from Wikipedia by using inter-language links for alignment. This  results in a large number of weakly comparable documents in more than $200$ languages. Wikipedia is a large source of multilingual data that is especially important for the languages for which no translation tools, multilingual dictionaries as Eurovoc \cite{eurovoc}, or strongly aligned multilingual corpora as Europarl \cite{europarl} are available. Documents in different languages are related with so called 'inter-language' links that can be found on the left of the Wikipedia page. The Wikipedia is constantly growing. There are currently 12 Wikipedias with more than 1 million %$10^6$
 articles, $52$ with more than 100k %$10^5$
 articles, $129$ with more than 10k articles, and $236$ with more than $1000$ articles.

Each Wikipedia page is embedded in the page tag. First, we check if the title of the page consists of any special namespace and do not process such pages. Then, we check if this is a redirection page and we store the redirect link because inter-language links can point to redirection links also. If none of the above applies, we extract the text and parse the Wikipedia markup. Currently, all the markup is removed.

@@

We observe that the method performs well on all pairs of languages, where at least 50,000 training documents are available(\emph{en}, \emph{es}, \emph{ru}, \emph{sl}). We note that taking $k = 500$ or $k = 1000$ multilingual topics usually results in similar performance, with some notable exceptions: in the case of (\emph{ht}, \emph{war}) the additional topics result in an increase in performance, as opposed to (\emph{ht},\emph{pms}) where performance drops, which suggests overfitting. The languages where the method performs poorly are \emph{ht} and \emph{war}, which can be explained by the quality of data (see Table \ref{table:rank} and explanation that follows). In case of \emph{pms}, we demonstrate that solid performance can be achieved for language pairs (\emph{pms}, \emph{sl}) and (\emph{pms}, \emph{hi}), where only 2000 training documents are shared between \emph{pms} and \emph{sl} and no training documents are available between \emph{pms} and \emph{hi}. Also observe that in the case of (\emph{pms}, \emph{ht}) the method still obtains a score of 62, even though training set intersection is zero and \emph{ht} data is corrupted, which we will show in the next paragraph.
{
\renewcommand\tabcolsep{3pt}
\begin{table}[h!]
\caption{Pairwise retrieval, 500 topics on the left -- 1000 topics on the right}\label{table:retrieval}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\cline{1-9}
&	en&	es&	ru&	sl&	hi&	war&	ht&	pms\\\cline{1-9}
en&	    &	98~-~98&	95~-~97&	97~-~98&	82~-~84&	76~-~74&	53~-~55&	 96~-~97\\
\cline{1-9}
es&	97~-~98&	&	94~-~96&	97~-~98&	85~-~84&	76~-~77&	56~-~57&	96~-~96\\
\cline{1-9}
ru&	96~-~97&	94~-~95&	&	97~-~97&	81~-~82&	73~-~74&	55~-~56&	96~-~96\\
\cline{1-9}
sl&	96~-~97&	95~-~95&	95~-~95&	&	91~-~91&	68~-~68&	59~-~69&	93~-~93\\
\cline{1-9}
hi&	81~-~82&	82~-~81&	80~-~80&	91~-~91&	&	68~-~67&	50~-~55&	87~-~86\\
\cline{1-9}
war&	68~-~63&	71~-~68&	72~-~71&	68~-~68&	66~-~62&	&	28~-~48&	 24~-~21\\
\cline{1-9}
ht&	52~-~58&	63~-~66&	66~-~62&	61~-~71&	44~-~55&	16~-~50&	&	62~-~49\\
\cline{1-9}
pms&	95~-~96&	96~-~96&	94~-~94&	93~-~93&	85~-~85&	23~-~26&	66~-~54&	 \\
\cline{1-9}
\end{tabular}
\end{center}
\end{table}
}

@@

% AMMR experiments
We also compared the proposed approaches on the task of Wikipedia mate retrieval (the same task as in Section~\ref{experiments:hubcca}). We computed the Average (over language pairs) Mean Reciprocal Rank (AMRR) \shortcite{voorhees1999trec}  performance of the different approaches on the  Wikipedia data by holding out $15,000$ aligned test documents and using $300,000$ aligned documents as the training set. The Figure~\ref{pic:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that hub CCA outperforms LSI approach and K-means lags far behind when testing on Wikipedia data. The hub CCA approach with $500$ topic vectors manages to perform comparably to the $LSI$-based approach with $1000$ topic vectors, which shows that the $CCA$ method can improve both model memory footprint as well as similarity computation time.

% number of topics?
Furthermore, we inspected how the number of topics influences the accuracy of cluster linking. As we can see from Table~\ref{table:linkingEvalAlgos} choosing number of topic vectors larger than $500$ barely affects linking performance, which is in contrast with the fact that additional topics helped to improve AMMR, see Figure~\ref{pic:AMRR}. Such differences may have arisen due to different domains of training and testing (Wikipedia pages versus news articles).

@@

%AMPR as function of the number of feature vectors
%500 is good, 800 better, 1000 overfitting, noise!
\begin{figure}
%\includegraphics{slika.png}
\centering
\input{retrieval}
\caption{Average of mean reciprocal ranks}
\label{pic:AMRR}
\end{figure}

\begin{table}[h]
\caption{Accuracy of cluster linking with 500 - 800 - 1000 topic vectors obtained from different cross-lingual similarity algorithms. The table shows for each of the algorithms the obtained classification accuracy, precision and recall.}
\label{table:linkingEvalAlgos}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  \cline{1-4}
  Models & A & Precision & Recall \\ \cline{1-4}
  CCA      & 78.2 - 79.6 - 80.3 \% & 76.3 - 78.0 - 80.5 \%  & 81.6 - 82.1 - 79.9 \%  \\ \cline{1-4}
  LSI      & 78.9 - 78.7 - 80.6 \%  & 76.8 - 77.0 - 78.7 \%  & 83.3 - 80.6 - 83.6 \%  \\ \cline{1-4}
 $k$-means & 73.9\phantom{ - 78.7 - 80.6 }\% & 69.5\phantom{ - 78.7 - 80.6 }\%  & 84.6\phantom{ - 78.7 - 80.6 }\%  \\ \cline{1-4}
\end{tabular}
\end{center}
\end{table}

@@

One of the main advantages of our approach is that it is highly scalable. It is fast, very robust to quality of training data, easily extendable, simple to implement and has relatively small hardware requirements. The similarity pipeline is the most computationally intensive part and currently runs on two Intel Xeon E5-2667 v2, 3.30GHz processor machine with 256GB of RAM. This is sufficient to do similarity computation over 150 languages if needed, see \url{http://xling.ijs.si} for live demo over 100 languages. It currently uses Wikipedia as a freely available knowledge base and experiments show that the similarity pipeline dramatically reduces the search space when linking clusters.

>>>>

The paper is organized as follows: we first provide an overview of the system as a whole (Section ~\ref{sec:pipeline}), followed by details on the cross-lingual similarity computation and our approach to linking clusters. We first introduce the problem of cross-lingual similarity computation (Section~\ref{sec:crosslingual}), which is followed by several approaches for measuring cross-lingual similarity (Section~\ref{sec:models}) and then the algorithms for matching clusters in different languages (Section~\ref{sec:linking}).  To improve readability, we will describe relevant related work for each module in its corresponding section. Finally, we present and interpret the experimental results, give conclusions and describe promising future directions.

@@

We base our techniques of cross-lingual event linking on an online system for detection of world events, called Event Registry \shortcite{Leban2014W,Leban2014I}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles that are collected from numerous news outlets all over the world. The important components in the pipeline of the Event Registry are shown in Figure~\ref{fig:erpipeline}. We will now briefly describe the main components.

@@

\subsection{Cross-Lingual Models}\label{sec:models}
In this section, we will describe several approaches to the problem of computing multilingual simliarities introduced in \ref{sec:tfidf}. We present four approaches:
a simple approach based on $k$-means clustering in \ref{sec:kmeans}, a standard approach based on singular value decomposition in \ref{sec:LSI}, a related
approach called Canonical Correlation Analysis (CCA) in \ref{sec:CCA} and finally a new method, which is an extension of CCA to more than two languages in \ref{sec:hublang}.
%
CCA can be used to find correlated patterns for a pair of languages, whereas the extended method optimizes a
Sum of Squared Correlations (SSCOR) between several language pairs, which was introduced in \cite{Kettenring}. The SSCOR problem is difficult to solve in our setting (hundreds of thousands of features, hundreds of thousands of examples). To tackle this, we propose a method which consists of two ingredients.
 The first one is based on an observation that certain datasets (such as Wikipedia) are biased towards one language (English), which can be exploited
 to reformulate a difficult optimization problem as an eigenvector problem. The second ingredient is dimensionality reduction using LSI, which
 makes the eigenvector problem computationally and numerically tractable.

@@

The resulting matrix (when appropriately scaled to have unit norm) for similarity computation between language $i$ and language $j$ is defined as (see the reasoning in the next paragraph):
$$C_i(C_i^T C_i)^{-1} (C_j^T C_j)^{-1} C_j.$$

@@

The first step in our method is to project $X_1, \ldots, X_m$ to lower-dimensional spaces without destroying the cross-lingual structure. Treating the nonzero columns of $X_i$ as observation vectors sampled from an underlying distribution $\mathcal{X}_i \in V_i = \RR^{n_i}$, we can analyze the empirical cross-covariance matrices:
$$C_{i,j} = \frac{1}{|a(i,j)|-1 }\sum_{\ell \in a(i,j)} (X_i^{\ell} - c_i)\cdot (X_j^{\ell} - c_j)^T,$$
 where $c_i = \frac{1}{a_i} \sum_{\ell \in a(i)}X_i^{\ell}$. By finding low-rank approximations of $C_{i,j}$ we can identify the subspaces of $V_i$ and $V_j$ that are relevant for extracting linear patterns between $\mathcal{X}_i$ and $\mathcal{X}_j$. Let $X_1$ represent the hub language corpus matrix. The LSI approach to finding the subspaces is to perform the singular value decomposition on the full $N \times N$ covariance matrix composed of blocks $C_{i,j}$. If $|a(i,j)|$ is small for many language pairs (as it is in the case of Wikipedia), then many empirical estimates $C_{i,j}$ are unreliable, which can result in overfitting. For this reason, we perform the truncated singular value decomposition on the matrix $C = [C_{1,2}  \cdots  C_{1,m}] \approx U S V^T$, where $U \in \RR^{n_1 \times k}, S \in \RR^{k \times k}, V \in \RR^{(\sum_{i=2}^m n_i) \times k}$. We split the matrix $V$ vertically in blocks with $n_2, \ldots, n_m$ rows: $V = [V_2^T  \cdots  V_m^T]^T$. Note that columns of $U$ are orthogonal but columns in each $V_i$ are not (columns of V are orthogonal). Let $V_1 := U$. We proceed by reducing the dimensionality of each $X_i$ by setting: $Y_i = V_i^T \cdot X_i$, where $Y_i \in \RR^{k\times N}$. To summarize, the first step reduces the dimensionality of the data and is based on CL-LSI, but optimizes only the hub language related cross-covariance blocks.

@@

We will describe the main dataset for building cross-lingual models which is based on Wikipedia and then present two sets of experiments. The first set of experiments
establishes that the hub based approach can deal with language pairs where little or no training data is available. The second set of experiments compares the main approaches
that we presented on the task of mate retrieval and the task of event linking. Finally, we examine how different choices of features impact the event linking performance.

@@

To investigate the empirical performance of the low-rank approximations we will test the algorithms on a large-scale, real-world multilingual dataset that we extracted from Wikipedia by using inter-language links for alignment. This  results in a large number of weakly comparable documents in more than $200$ languages. Wikipedia is a large source of multilingual data that is especially important for the languages for which no translation tools, multilingual dictionaries as Eurovoc \cite{eurovoc}, or strongly aligned multilingual corpora as Europarl \cite{europarl} are available. Documents in different languages are related with so called 'inter-language' links that can be found on the left of the Wikipedia page. The Wikipedia is constantly growing. There are currently 12 Wikipedias with more than 1 million %$10^6$
 articles, $52$ with more than 100k %$10^5$
 articles, $129$ with more than 10k articles, and $236$ with more than $1,000$ articles.

Each Wikipedia page is embedded in the page tag. First, we check if the title of the page consists of any special namespace (which includes categories and discussion pages) and do not process such pages. Then, we check if this is a redirection page and we store the redirect link because inter-language links can point to redirection links also. If none of the above applies, we extract the text and parse the Wikipedia markup. Currently, all the markup is removed.

@@

We observe that the method performs well on all pairs of languages, where at least 50,000 training documents are available(\emph{en}, \emph{es}, \emph{ru}, \emph{sl}). We note that taking $k = 500$ or $k = 1,000$ multilingual topics usually results in similar performance, with some notable exceptions: in the case of (\emph{ht}, \emph{war}) the additional topics result in an increase in performance, as opposed to (\emph{ht},\emph{pms}) where performance drops, which suggests overfitting. The languages where the method performs poorly are \emph{ht} and \emph{war}, which can be explained by the quality of data (see Table \ref{table:rank} and explanation that follows). In case of \emph{pms}, we demonstrate that solid performance can be achieved for language pairs (\emph{pms}, \emph{sl}) and (\emph{pms}, \emph{hi}), where only 2,000 training documents are shared between \emph{pms} and \emph{sl} and no training documents are available between \emph{pms} and \emph{hi}. Also observe that in the case of (\emph{pms}, \emph{ht}) the method still obtains a score of 62, even though training set intersection is zero and \emph{ht} data is corrupted, which we will show in the next paragraph.
{
\renewcommand\tabcolsep{3pt}
\begin{table}[h!]
\caption{Pairwise retrieval, 500 topics on the left -- 1,000 topics on the right}\label{table:retrieval}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\cline{1-9}
&	en&	es&	ru&	sl&	hi&	war&	ht&	pms\\\cline{1-9}
en&	    &	98~-~98&	95~-~97&	97~-~98&	82~-~84&	76~-~74&	53~-~55&	 96~-~97\\
\cline{1-9}
es&	97~-~98&	&	94~-~96&	97~-~98&	85~-~84&	76~-~77&	56~-~57&	96~-~96\\
\cline{1-9}
ru&	96~-~97&	94~-~95&	&	97~-~97&	81~-~82&	73~-~74&	55~-~56&	96~-~96\\
\cline{1-9}
sl&	96~-~97&	95~-~95&	95~-~95&	&	91~-~91&	68~-~68&	59~-~69&	93~-~93\\
\cline{1-9}
hi&	81~-~82&	82~-~81&	80~-~80&	91~-~91&	&	68~-~67&	50~-~55&	87~-~86\\
\cline{1-9}
war&	68~-~63&	71~-~68&	72~-~71&	68~-~68&	66~-~62&	&	28~-~48&	 24~-~21\\
\cline{1-9}
ht&	52~-~58&	63~-~66&	66~-~62&	61~-~71&	44~-~55&	16~-~50&	&	62~-~49\\
\cline{1-9}
pms&	95~-~96&	96~-~96&	94~-~94&	93~-~93&	85~-~85&	23~-~26&	66~-~54&	 \\
\cline{1-9}
\end{tabular}
\end{center}
\end{table}
}

@@

% AMMR experiments
We also compared the proposed approaches on the task of Wikipedia mate retrieval (the same task as in Section~\ref{experiments:hubcca}). We computed the Average (over language pairs) Mean Reciprocal Rank (AMRR) \shortcite{voorhees1999trec}  performance of the different approaches on the  Wikipedia data by holding out $15,000$ aligned test documents and using $300,000$ aligned documents as the training set. The Figure~\ref{pic:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that hub CCA outperforms LSI approach and K-means lags far behind when testing on Wikipedia data. The hub CCA approach with $500$ topic vectors manages to perform comparably to the $LSI$-based approach with $1,000$ topic vectors, which shows that the $CCA$ method can improve both model memory footprint as well as similarity computation time.

% number of topics?
Furthermore, we inspected how the number of topics influences the accuracy of cluster linking. As we can see from Table~\ref{table:linkingEvalAlgos} choosing a number of features larger than $500$ barely affects linking performance, which is in contrast with the fact that additional topics helped to improve AMMR, see Figure~\ref{pic:AMRR}. Such differences may have arisen due to different domains of training and testing (Wikipedia pages versus news articles).

@@

%AMPR as function of the number of feature vectors
\begin{figure}
%\includegraphics{slika.png}
\centering
\input{retrieval}
\caption{Average of mean reciprocal ranks}
\label{pic:AMRR}
\end{figure}

\begin{table}[h]
\caption{Accuracy of cluster linking with 500 - 800 - 1,000 topic vectors obtained from different cross-lingual similarity algorithms. The table shows for each of the algorithms the obtained classification accuracy, precision and recall.}
\label{table:linkingEvalAlgos}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  \cline{1-4}
  Models & A & Precision & Recall \\ \cline{1-4}
  CCA      & 78.2 - 79.6 - 80.3 \% & 76.3 - 78.0 - 80.5 \%  & 81.6 - 82.1 - 79.9 \%  \\ \cline{1-4}
  LSI      & 78.9 - 78.7 - 80.6 \%  & 76.8 - 77.0 - 78.7 \%  & 83.3 - 80.6 - 83.6 \%  \\ \cline{1-4}
 $k$-means & 73.9\phantom{ - 78.7 - 80.6 }\% & 69.5\phantom{ - 78.7 - 80.6 }\%  & 84.6\phantom{ - 78.7 - 80.6 }\%  \\ \cline{1-4}
\end{tabular}
\end{center}
\end{table}

@@

One of the main advantages of our approach is that it is highly scalable. It is fast, very robust to quality of training data, easily extendable, simple to implement and has relatively small hardware requirements. The similarity pipeline is the most computationally intensive part and currently runs on a machine with two Intel Xeon E5-2667 v2, 3.30GHz processors with 256GB of RAM. This is sufficient to do similarity computation over a large number of languages if needed. It currently uses Wikipedia as a freely available knowledge base and experiments show that the similarity pipeline dramatically reduces the search space when linking clusters.
